### 2016-10-18 capability-machine.tex
Updated structure of introduction of the logical relation (still work in progress, so the same thing is repeated), so it was easier to prove well-foundedness and the existense of the value relation.

%Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\finparfun}{\stackrel{\textit{\tiny{fin}}}{\rightharpoonup}}
\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\longrightarrow}}
\newcommand{\fun}{\longrightarrow}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}
\newcommand{\blater}{\mathop{\blacktriangleright}}
\newcommand{\id}{\var{id}}
\newcommand{\undefined}{\mathit{undefined}}

@@

\newcommand{\Rel}{\plaindom{Rel}}
\newcommand{\States}{\plaindom{State}}
\newcommand{\RegionNames}{\plaindom{RegionName}}
\newcommand{\Regions}{\plaindom{Region}}
\newcommand{\Worlds}{\plaindom{World}}
\newcommand{\Wor}{\plaindom{Wor}}
\newcommand{\StorePred}{\plaindom{HeapSegPred}}
\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}

@@

  \updatePcPerm{(\perm,\start,\addrend,\addr)} &=

@@

                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}\\
                                       (\perm,\start,\addrend,\addr) & \text{otherwise} 

@@

@@

We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
work use \dominique{this sentence doesn't parse :)}
\begin{definition}[Preordered \cofe{}\ construction: Function]
  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
\[
  \begin{gathered}
    f \future g \\
    \text{ iff } \\
    \forall s \in \dom(f) \ldotp f(s) \future g(s)
  \end{gathered}
\]
\end{definition}

The above are standard constructions, so they are used here without showing they
are in fact well-defined.

\dominique{What is HS? Do you mean $\HeapSegments$?}
\dominique{Why the exponential in the definition of $G$, why not $W^{\States} \monnefun \UPred{HS}$?}
Define functors $K$, $R$, and $G$ as follows:
\begin{align*}
  & K : \CatP \rightarrow \CatP \\
  & K(R) = \nats \finparfun R \\
  & K(f) = \lambda \phi \ldotp \lambda n \ldotp f(\phi(n)) \\
  & \\
  & R : \CatC \rightarrow \CatP \\
  & R(H) = \States \times \Rel \times H\\
  & R(h) = \lambda (s,\Phi,H) \ldotp (s,\Phi,h(H))\\
  & \\
  & G : \CatP^{\var{op}} \rightarrow \CatC \\
  & G(W) = \UPred{HS}^{W^{\States}} \\
  & G(g) = \lambda H \ldotp \lambda \var{st} \ldotp \lambda x \ldotp H (\var{st})(g(x))
\end{align*}
We first show that $K$, $R$, and $G$ are well-defined mappings.
\begin{lemma}[World finite partial mapping]
  \label{lem:world-finpar}
  For all $f$ and $\phi$, $K(f)(\phi)$ is a finite partial mapping.
\end{lemma}
\begin{lemma}[Heap segment predicate monotone]
\label{lem:HSP-mon}
For all $g$, $H$, and $\var{st}$
\[
G(g) (H) (\var{st})
\]
is non-expansive.
\end{lemma}
\begin{lemma}[Heap segment predicate non-expansive]
\label{lem:HSP-ne}
For all $g$, $H$, and $\var{st}$
\[
G(g) (H) (\var{st})
\]
is monotone.
\end{lemma}
Next we show that $K$, $R$, and $G$ are in fact functors:
\begin{lemma}[$K$ functorial]\forcenewline
\label{lem:K-func}
  \begin{enumerate}
  \item $K(f) : K(X) \rightarrow K(Y)$ is monotone and non-expansive for $f : X \monnefun Y$
  \item $K(f \circ g) = K(f) \circ K(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
  \item $K(\id) = \id$
  \end{enumerate}
\end{lemma}

@@

\subsection{Worlds}
Assume preordered \cofe{}\ $\Wor$ and isomorphism $\xi$ such that:
\[
  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
\]
We now define regions as
\[
\Regions \defeq (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
\]
and we define worlds as
\[
\Worlds \defeq \RegionNames \finparfun \Regions
\]
To define future worlds and regions, We use the ordering inherited from the preordered \cofes{}.
\begin{definition}[Future worlds]
  For $W, W' \in \Worlds$
 \begin{align*}
 W' \future W & & \text{ iff } & &
   \begin{gathered}
     \dom(W') \supseteq \dom(W)\\ 
     \text{ and } \\
     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
   \end{gathered}
 \end{align*}
\end{definition}

@@

\subsection{Standard Regions}
\label{subsec:standard-regions}
A standard heap invariant that insures all values in the region are in the value  relation:
\[
  \iota_{\var{start},\var{end}} \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
  H_{\var{std}}\; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
                                               \begin{aligned}
                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
                                               \end{aligned} \right\}
\end{align*}
\begin{lemma}[$\iota_{\var{start},\var{end}}$ is well-defined]
\label{lem:std-region-wd}
  For all $\var{base}$ and $\var{end}$, $H_{\var{std}}\; (\var{base}, \var{end})$ is monotone and non-expansive. $=$ is a reflexive and transitive relation.
\end{lemma}

>>>>

%Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\finparfun}{\stackrel{\textit{\tiny{fin}}}{\rightharpoonup}}
\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\rightarrow}}
\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\rightarrow}}
\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\rightarrow}}
\newcommand{\fun}{\rightarrow}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}
\newcommand{\blater}{\mathop{\blacktriangleright}}
\newcommand{\id}{\var{id}}
\newcommand{\undefined}{\mathit{undefined}}

@@

\newcommand{\Rel}{\plaindom{Rel}}
\newcommand{\States}{\plaindom{State}}
\newcommand{\RegionNames}{\plaindom{RegionName}}
\newcommand{\Regions}{\plaindom{Region}}
\newcommand{\Worlds}{\plaindom{World}}
\newcommand{\Wor}{\plaindom{Wor}}
\newcommand{\StorePred}{\plaindom{HeapSegPred}}
\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
\newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}

@@

  \updatePcPerm{w} &=

@@

                                       (\exec,\start,\addrend,\addr) & \text{if $w = (\entry,\start,\addrend,\addr)$}\\
                                       w & \text{otherwise} 

@@

We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
\dominique{this sentence doesn't parse :)}
\begin{definition}[Preordered \cofe{}\ construction: Function]
  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
\[
  \begin{gathered}
    f \future g \\
    \text{ iff } \\
    \forall s \in \dom(f) \ldotp f(s) \future g(s)
  \end{gathered}
\]
\end{definition}

The above are standard constructions, so they are used here without showing they
are in fact well-defined as shown in \cite{Birkedal:tutorial-notes}.

@@

Define functors $K$, $R$, and $G$ as follows:
\begin{align*}
  & K : \CatP \rightarrow \CatP \\
  & K(R) = \nats \finparfun R \\
  & K(f) = \lambda \phi \ldotp \lambda n \ldotp f(\phi(n)) \\
  & \\
  & R : \CatC \rightarrow \CatP \\
  & R(H) = \States \times \Rel \times H\\
  & R(h) = \lambda (s,\Phi,H) \ldotp (s,\Phi,h(H))\\
  & \\
  & G : \CatP^{\var{op}} \rightarrow \CatC \\
  & G(W) = \States \nefun W \monnefun \UPred{HS} \\
  & G(g) = \lambda H \ldotp \lambda \var{st} \ldotp \lambda x \ldotp H (\var{st})(g(x))
\end{align*}
We first show that $K$, $R$, and $G$ are well-defined mappings.
\begin{lemma}[World finite partial mapping]
  \label{lem:world-finpar}
  For all $f$ and $\phi$, $K(f)(\phi)$ is a finite partial mapping.
\end{lemma}
\begin{lemma}[Heap segment predicate monotone]
\label{lem:HSP-mon}
For all $g$, $H$, and $\var{st}$
\[
G(g) (H) (\var{st})
\]
is non-expansive.
\end{lemma}
\begin{lemma}[Heap segment predicate non-expansive]
\label{lem:HSP-ne}
For all $g$, $H$, and $\var{st}$
\[
G(g) (H) (\var{st})
\]
is monotone.
\end{lemma}
Next we show that $K$, $R$, and $G$ are in fact functors:
\begin{lemma}[$K$ functorial]\forcenewline
\label{lem:K-func}
  \begin{enumerate}
  \item $K(f) : K(X) \rightarrow K(Y)$ is monotone and non-expansive for $f : X \monnefun Y$
  \item $K(f \circ g) = K(f) \circ K(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
  \item $K(\id) = \id$
  \end{enumerate}
\end{lemma}

@@

\subsection{Worlds}
Assume preordered \cofe{}\ $\Wor$ and isomorphism $\xi$ such that:
\[
  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
\]
We now define regions as
\[
\Regions \defeq (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
\]
define region names to be natural numbers, i.e.,
\[
  \RegionNames \defeq \nats
\]
and define worlds as
\[
\Worlds \defeq \RegionNames \finparfun \Regions
\]
To define future worlds and regions, We use the ordering inherited from the preordered \cofes{}.
\begin{definition}[Future worlds]
  For $W, W' \in \Worlds$
 \begin{align*}
 W' \future W & & \text{ iff } & &
   \begin{gathered}
     \dom(W') \supseteq \dom(W)\\ 
     \text{ and } \\
     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
   \end{gathered}
 \end{align*}
\end{definition}

@@

\subsection{Logical Relation}
\label{subsec:logical-relation}
Our logical relation is defined using multiple recursive definitions, so the definitions in the following subsections are defined simultaneously. We want to define the value relation as the fixed-point given by Banach's fixed-point theorem, so all our definitions will be parameterized with the value relation.
\subsubsection{Observation Relation}
\label{subsubsec:observation-relation}
In order to define the expression relation, we define an observation relation.
\begin{align*}
  \observations \quad : & \quad  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
\end{align*}
A pair of a register and a heap segment is ``good'' if we can put it together with a frame heap, so we can execute it. The execution should then end up in a heap where the frame remains the same and the remaining heap segment satisfies the world.

Note that the operational semantic is total, so we cannot get stuck. If the execution ends up in a $\failed$ configuration, then we do not care about the heap and the registers. This is why, we only have requirements on the result when we end up in a $\halted$ configuration.

The following lemmas show that the observation relation is well-defined.
\begin{lemma}[Observation relation uniformity]
\label{lem:obs-dc}
  \begin{align*}
    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp\\
    &\quad \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
  \end{align*}
\end{lemma}

\begin{lemma}[Observation relation non-expansive in worlds]
\label{lem:obs-ne-worlds}
\begin{align*}
  &\forall W, W', n \ldotp \\
  &\quad W \nequal W' \Rightarrow \observations(W) \nequal \observations(W')
\end{align*}
\end{lemma}

\subsubsection{Regiser-File Relation}
\label{subsubsec:register-file-relation}
This relation is used in the definition of the continuation relation as well as the expression relation.
\begin{align*}
  \stdrr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Regs} \\
  \stdrr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp
                     \begin{aligned}[t]
                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
                     \end{aligned}
\end{align*}
Well-formedness lemmas for this definition:
\begin{lemma}[Register relation uniformity]
\label{lem:reg-dc}
  \begin{align*}
    &\forall \stdvr, n' < n \ldotp \forall W \ldotp \forall \reg \ldotp\\
    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair[n']{\reg} \in \stdrr(\stdvr)(W)
  \end{align*}
\end{lemma}

\begin{lemma}[Register relation montone in worlds]
\label{lem:reg-mono-worlds}
  \begin{align*}
    &\forall \stdvr, n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp\\
    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair{\reg} \in \stdrr(\stdvr)(W')
  \end{align*}
\end{lemma}

\begin{lemma}[Register relation non-expansive in value relation]
\label{lem:reg-ne-vr}
\begin{align*}
  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdrr(\stdvr) \nequal \stdrr(\stdvr')
\end{align*}
\end{lemma}

\subsubsection{Continuation Relation}
\label{subsubsection:continuation-relation}
The continuation relation is used in the definition of the expression relation. The continuation relation ensures that if you continue execution through a continuation, then it will result in a good result according to the world.
\begin{align*}
  \stdkr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
                     \begin{aligned}[t]
                       & \npair{c} \in \stdvr(W) \land \\
                       &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(\stdvr)(W') \ldotp \\
                       & \quad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
                 \end{aligned}
\end{align*}

Well-definedness lemmas:
\begin{lemma}[Continuation relation uniformity]
\label{lem:stdkr-dc}
  \begin{align*}
    &\forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp\\
    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair[n']{c} \in \stdkr(\stdvr)(W)
  \end{align*}
\end{lemma}

\begin{lemma}[Continuation relation monotone in worlds]
\label{lem:stdkr-mono-worlds}
  \begin{align*}
    &\forall \stdvr \ldotp \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp\\
    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair{c} \in \stdkr(\stdvr)(W')
  \end{align*}
\end{lemma}

\begin{lemma}[Continuation relation non-expansive in value relation]
\label{lem:stdkr-ne-vr}
\begin{align*}
  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdkr(\stdvr) \nequal \stdkr(\stdvr')
\end{align*}
\end{lemma}

\subsubsection{Expression Relation}
\label{subsubsection:expression-relation}
The expression relation is defined as follows:
\begin{align*}
  \stder \quad : & \quad  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \nefun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
  \stder \defeq & \lambda \stdvr \ldotp \lambda W \ldotp \{ \npair{\pc} \mid 
  \begin{aligned}[t]
    & \forall n' \leq n \ldotp\\
    & \quad \forall \npair[n']{\reg} \in \stdrr(\stdvr)(W) \ldotp \\
    & \quad \quad \forall \npair[n']{c} \in \stdkr(\stdvr)(W) \ldotp \\
    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
  \end{aligned}
\end{align*}
Well-definedness lemmas:
\begin{lemma}[Expression relation uniformity]
\label{lem:stder-dc}
  \begin{align*}
    & \forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp\\
    & \quad \npair{\pc} \in \stder(\stdvr)(W) \Rightarrow \npair[n']{\pc} \in \stder(\stdvr)(W) 
 \end{align*}
\end{lemma}
\begin{lemma}[Expression relation non-expansive in world]
\label{lem:stder-ne-worlds}
\begin{align*}
  \forall \stdvr \ldotp \forall W_1 \nequal W_2 \ldotp \stder(\stdvr)(W_1) \nequal \stder(\stdvr)(W_2)
\end{align*}
\end{lemma}

\begin{lemma}[Expression relation non-expansive in value relation]
\label{lem:stder-ne-vr}
\begin{align*}
  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stder(\stdvr) \nequal \stder(\stdvr')
\end{align*}
\end{lemma}

\subsubsection{Standard Region}
\label{subsubsec:standard-region}
The following standard region is used in the definition of the value relation. Specifically, it is used in the $\readCond{}$ and the $\writeCond{}$ (to be defined next)

\begin{align*}
  \iota_{\var{start},\var{end}} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun \Regions\\
  \iota_{\var{base},\var{end}} & \defeq \lambda \; \stdvr \ldotp ((\var{base},\var{end}),=,H_{\var{std}} (\stdvr)) \\ & \\
  H_{\var{std}} & : (\Worlds \monnefun \UPred{\Words}) \nefun \States \nefun \Worlds \monnefun \UPred{\HeapSegments} \\
  H_{\var{std}} \; \stdvr \; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
                                               \begin{aligned}
                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
                                               \end{aligned} \right\}
\end{align*}
As mentioned previously, the set of states contains the ``necessary'' states. For the above to make sense, the set of states contains pairs of natural numbers $(\var{base},\var{end})$.

The well-definedness lemmas for the above is:
\begin{lemma}[$H_{\var{std}}$ is monotone in the worlds]
\label{lem:std-reg-mono-world}
\begin{align*}
  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall W' \future W \ldotp \\
  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W' \supseteq H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
\end{align*}
\end{lemma}

\begin{lemma}[$H_{\var{std}}$ is non-expansive in the worlds]
\label{lem:std-reg-ne-world}
\begin{align*}
  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall n \ldotp \forall W_1 \nequal W_2 \ldotp \\
  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W_1 \nequal H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
\end{align*}
\end{lemma}

\begin{lemma}[$H_{\var{std}}$ is non-expansive in the value relation]
\label{lem:std-reg-ne-vr}
\begin{align*}
  & \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
  & \quad H_{\var{std}}\; \stdvr \nequal H_{\var{std}}\; \stdvr'
\end{align*}
\end{lemma}

\begin{lemma}[$\iota_{\start,\addrend}$ is non-expansive in the value relation]
\label{lem:std-reg-i-ne-vr}
\begin{align*}
  & \forall \start, \addrend \ldotp \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
  & \quad \iota_{\start,\addrend}\; \stdvr \nequal \iota_{\start,\addrend}\; \stdvr'
\end{align*}
\end{lemma}

\subsubsection{Capability Conditions}
\label{subsubsec:capability-conditions}
The definition of the value relation has the same conditions several times, so to define it consisely, we define the following conditions.
\begin{align*}
  \readCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
  \readCond{}(\stdvr)(\start,\addrend,W) & =        
  \begin{aligned}[t]
    \{ n \mid & \;\exists r \in \RegionNames \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(\stdvr) \}
  \end{aligned}\\
 & \\
  \writeCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
  \writeCond{}(\stdvr)(\start,\addrend, W) & = 
  \begin{aligned}[t]
    \{ n \mid & \; \exists r \in \RegionNames \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \; \qquad W(r) \nequal[n-1] \iota_{\start',\addrend'}(\stdvr)  \}
  \end{aligned}\\
 & \\
  \execCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Perms \times \Worlds) \monnefun \DCPred{\nats} \\
  \execCond{}(\stdvr)(\start,\addrend,\perm,W) & = 
  \begin{aligned}[t]
    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
              & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
              & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
  \end{aligned} \\
 & \\
  \entryCond{} & : \;  (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^3 \times \Worlds) \monnefun \DCPred{\nats} \\
  \entryCond{}(\stdvr)(\start,\addrend,\addr,W) & =
  \begin{aligned}[t]
    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
              & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
  \end{aligned}
\end{align*}

The following lemmas show that the above conditions are well-defined:
\begin{lemma}[Read condition downwards-closed]
\label{lem:rc-dc}
  \begin{align*}
    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
    &\quad  n \in \readCond{}(\stdvr)(\start,\addrend,W) \land \\
    &\quad  n' \leq n \\
    &\qquad \Rightarrow n' \in \readCond{}(\stdvr)(\start,\addrend,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Read condition monotone in world]
\label{lem:rc-mono-world}
  \begin{align*}
    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
    &\qquad \Rightarrow \readCond{}(\stdvr)(\start,\addrend,W') \supseteq \readCond{}(\stdvr)(\start,\addrend,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Read condition non-expansive in worlds]
\label{lem:rc-ne-world}
  \begin{align*}
    & \forall \stdvr, b,e, n, W, W' \ldotp \\
    & \quad (b,e,W) \nequal (b,e,W') \Rightarrow \\
    & \qquad \readCond{}(\stdvr)(b,e,W) \nequal \readCond{}(\stdvr)(b,e,W')\\
  \end{align*}
\end{lemma}

\begin{lemma}[Read-write condition uniformity]
\label{lem:wc-dc}
  \begin{align*}
    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
    &\quad  n \in \writeCond{}(\stdvr)(\start,\addrend,W) \land \\
    &\quad  n' \leq n \\
    &\qquad \Rightarrow n' \in \writeCond{}(\stdvr)(\start,\addrend,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Read-write condition monotone in world]
\label{lem:wc-mono-world}
  \begin{align*}
    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start,\addrend,W') \supseteq \writeCond{}(\stdvr)(\start,\addrend,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Execute condition downwards-closed]
\label{lem:ec-dc}
  \begin{align*}
    &\forall \stdvr, n,n',W,\start, \addrend, \perm \ldotp \\
    &\quad  n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W) \land \\
    &\quad  n' \leq n \\
    &\qquad \Rightarrow n' \in \execCond{}(\stdvr)(\start,\addrend,\perm,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Execute condition monotone in world]
\label{lem:ec-mono-world}
  \begin{align*}
    &\forall \stdvr,n,W, W',\start, \addrend, \perm \ldotp \\
    &\quad  (\start,\addrend,\perm,W') \future (\start,\addrend,\perm,W) \\
    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \execCond{}(\stdvr)(\start,\addrend,\perm,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Execute condition non-expansive in worlds]
\label{lem:ec-ne-world}
  \begin{align*}
    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\perm \ldotp \\
    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
  \end{align*}
\end{lemma}

\begin{lemma}[Entry condition downwards-closed]
\label{lem:en-dc}
  \begin{align*}
    &\forall \stdvr, n,n',W,\start, \addrend, \addr \ldotp \\
    &\quad  n \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W) \land \\
    &\quad  n' \leq n \\
    &\qquad \Rightarrow n' \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Entry condition monotone in world]
  \label{lem:en-mono-world}
  \begin{align*}
    &\forall \stdvr,n,W, W',\start, \addrend, \addr \ldotp \\
    &\quad  (\start,\addrend,\addr,W') \future (\start,\addrend,\addr,W) \\
    &\qquad \Rightarrow \entryCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \entryCond{}(\stdvr)(\start,\addrend,\perm,W)
  \end{align*}
\end{lemma}

\begin{lemma}[Entry condition non-expansive in worlds]
\label{lem:en-ne-world}
  \begin{align*}
    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\addr \ldotp \\
    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
  \end{align*}
\end{lemma}
Finally, we need to show that all the conditions are non-expansive, but we later want to use Banach's fixed point theorem to define the value relation. For this we will need that the above conditions are contractive, and if they are contractive, then they are also non-expansive, so we show that each of the conditions are contractive:
\begin{lemma}[Read condition contractive]
  \label{lem:rc-con}
  \begin{align*}
    &\forall \stdvr, \stdvr', n \ldotp \\
    &\quad \stdvr \nequal \stdvr' \Rightarrow \readCond{}(\stdvr) \nequal[n+1] \readCond{}(\stdvr')
  \end{align*}
\end{lemma}

\begin{lemma}[Write condition contractive]
  \label{lem:wc-con}
  \begin{align*}
    &\forall \stdvr, \stdvr', n \ldotp \\
    &\quad \stdvr \nequal \stdvr' \Rightarrow \writeCond{}(\stdvr) \nequal[n+1] \writeCond{}(\stdvr')
  \end{align*}
\end{lemma}

\begin{lemma}[Execute condition contractive]
  \label{lem:ec-con}
  \begin{align*}
    &\forall \stdvr, \stdvr', n \ldotp \\
    &\quad \stdvr \nequal \stdvr' \Rightarrow \execCond{}(\stdvr) \nequal[n+1] \execCond{}(\stdvr')
  \end{align*}
\end{lemma}

\begin{lemma}[Entry condition contractive]
  \label{lem:en-con}
  \begin{align*}
    &\forall \stdvr, \stdvr', n \ldotp \\
    &\quad \stdvr \nequal \stdvr' \Rightarrow \entryCond{}(\stdvr) \nequal[n+1] \entryCond{}(\stdvr')
  \end{align*}
\end{lemma}

\subsubsection{Value Relation}
\label{subsubsec:value-relation}
The value relation, is defined as follows:
\begin{align*}
  \stdvr & :  (\Worlds \monnefun \UPred{Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
              \begin{aligned}[t]
  & \{ \npair{i} \mid i \in \ints \} 
    \union \\
         & \{ \npair{\stdcap[\noperm] }  \} 
           \union \\
         & \{ \npair{\stdcap[\readonly] } \mid n \in \readCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\readwrite] } \mid n \in \writeCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\exec]} \mid \\
         & \quad n \in \readCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \}
           \union \\
           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
         & \{ \npair{\stdcap[\entry]} \mid n \in \entryCond{}(V)(\start,\addrend,\addr,W) \} \union \\
         & \{ \npair{\stdcap[\rwx]} \mid \\
         & \quad n \in \writeCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\rwx,W) \}
       \end{aligned}
\end{align*}
